NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2020-09-16 Manuscript/manuscriptPGPE.tex

      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\

@@

      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  300  & 170 & 100 & \cite{POPCslipid300K} \\
      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\

@@

      POPC                   & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\

@@

\clearpage
\subsection{PC headgroup interactions with PE and PG}
\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
  \caption{\label{HGorderparametersPCvsPEPG}
    Modulation of POPC headgroup order parameters with increasing amount of POPE (left) and POPG (right) in bilayer
    from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  %\todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
\end{figure*}

@@

For $\beta$-carbon order parameter in PG headgroup, experiments report
mild increase \cite{macdonald87} or no change~\cite{borle85} upon addition 
of PC lipids (Fig. \ref{HGorderparametersPGvsPCchange}). 
Simulations with all the tested force fields give only very small changes also for
the $\alpha$-carbon order parameter (Figs. \ref{HGorderparametersPGvsPC} and \ref{HGorderparametersPGvsPCchange}). 
Therefore, the simulations are generally in line with experiments, suggesting that the
interactions with PC do not essentially effect the PG headgroup structure.
This suggests that the interactions between PG and PC headgroups are captured better in
simulations than for PS headgroup, where
all the force fieds significantly overestimated the structural response of PS headgroup
to the interactions with PC lipids \cite{antila18}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPCchange.eps}
  \caption{\label{HGorderparametersPGvsPCchange}
    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
  }
\end{figure}

@@

\clearpage
\subsection{Calcium binding to PE and PG lipid bilayers}
\begin{figure*}[bt]
  \centering
  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
  \caption{\label{changesWITHCaClPG}
    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
    The changes with respect to the systems without CaCl$_2$ are shown for other data than
    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
  }
  \todo{Data for CHARMM with 100mM is still coming.}
\end{figure*}

>>>>

%      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
%      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
      POPC                   & CHARMM36 \cite{??}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\

@@

      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  298  & 170 & 100 & \cite{POPCslipid298K} \\
      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120  & 0  &  298  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\

@@

      POPC                   & Slipid \cite{jambeck12b}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\

@@

\clearpage
\subsection{PC headgroup interactions with PE and PG}
\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
  \caption{\label{HGorderparametersPCvsPEPG}
    Modulation of POPC headgroup order parameters with increasing amount of POPE (left) and POPG (right) in bilayer
    from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{P-N vector angles from Berger-OPLS, Lipid17, and GROMOS-CKP simulations are yet to be analyzed.}
\end{figure*}

@@

For $\beta$-carbon order parameter in PG headgroup, experiments report
mild increase \cite{macdonald87} or no change~\cite{borle85} upon addition 
of PC lipids (Fig. \ref{HGorderparametersPGvsPCchange}). 
Simulations with all the tested force fields give only very small changes also for
the $\alpha$-carbon order parameter (Figs. \ref{HGorderparametersPGvsPC} and \ref{HGorderparametersPGvsPCchange}). 
Therefore, the simulations are generally in line with experiments, suggesting that the
interactions with PC do not essentially effect the PG headgroup structure.
This suggests that the interactions between PG and PC headgroups are captured better in
simulations than for PS headgroup, where
all the force fieds significantly overestimated the structural response of PS headgroup
to the interactions with PC lipids \cite{antila18}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPCchange.eps}
  \caption{\label{HGorderparametersPGvsPCchange}
    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
  }
  \todo{P-N angles from Lipid17, Slipids, and GROMOS-CKP yet to be calculated.}
\end{figure}

@@

\clearpage
\subsection{Calcium binding to PE and PG lipid bilayers}
\begin{figure*}[bt]
  \centering
  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
  \caption{\label{changesWITHCaClPG}
    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
    The changes with respect to the systems without CaCl$_2$ are shown for other data than
    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
  }
  \todo{Data for CHARMM with 100mM is still coming.} \\
  \todo{P-N vector angles from Lipid17 and Lipid17ecc models yet to be analyzed.}
\end{figure*}
